---
title: "Computational & Synthetic Biology"
author: "Alexa"
date: "2025-01-12"
output: html_document
---

## Data

Importing data from fireprot 2.0 job 4nudmg (<https://loschmidt.chemi.muni.cz/fireprotweb/job/4nudmq>)

```{r, data_import}
fireprot_data <- read_excel("./fireprot_data.xlsx", range = "B2:I14309")
```

## Potential Mutations in catalytic site

Positions marked as active in catalytic site have been marked in excel file beforehand. Could be interesting to modify to increase activity, however are likely highly conserved.

#### Check how conserved catalytic site enzymes are: 

```{r, catalytic}
library(dplyr)
fireprot_data %>% filter(`Catalytic site` == 1) %>% select(position, conservation_score) %>% table
```

Catalytic enzymes are highly conserved, with conservation scores ranging from 6 to 9.

S642 is the catalytic enzyme with the lowest conservation and therefore most suitable for mutation.

#### What mutations could be introduced at S642?

```{r, S642}
fireprot_data %>% filter(position == 642) %>% select(mutation, foldx, rosetta) 
```

No foldx or rosetta score available to guide choice of mutation. Choice by biological concepts, e.g. chose AA with properties similar to Serine.

## Potential Mutations outside of catalytic site

Efficiency of an enzyme can not only be improved by changing catalytic site but also by improving stability. Higher stability can for example allow function under a higher temperature which usually increases speed of reactions.

#### Mutations for stability improvement

Wanted:

-   low conservation score (changes less likely to disturb enzyme)

-   low foldx and rosetta score (show change of free energy after mutations, negative values implicate improved stability)

Filter for only those mutations with foldx + rosetta score available and determine the distribution of conservation scores

```{r}
fireprot_data %>% filter(foldx != "null" & rosetta != "null") %>% select(conservation_score) %>% table
```

Conservation scores low, further investigate mutations with 1 or 2 (artificial cut off).

```{r}
fireprot_data %>% filter(foldx != "null" & rosetta != "null" & conservation_score<= 2) %>% select(position, residue, mutation, foldx, rosetta, conservation_score) -> mutation_candidates

# apply right data type 
mutation_candidates %>% mutate(foldx = as.double(foldx), rosetta = as.double(rosetta)) -> mutation_candidates

print(mutation_candidates)
```

These 10 mutations could be experimentally tested for their effect on efficiency and stability.

```{r}
getwd()

```

